---
title             : "Manuscript"
shorttitle        : "Subtitle"

author: 
  - name          : "Sally E. Claridge"
    affiliation   : "1"
  - name          : "Daniel M. Charytonowicz"
    affiliation   : "1"
  - name          : "Adam A. Margolin"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "Icahn School of Medicine at Mount Sinai, Deparment of Genetics and Genomic Sciences, One Gustave L. Levy Place, Box 1498, New York, NY 10029"
    email         : "adam.margolin@mssm.edu"

affiliation:
  - id            : "1"
    institution   : "Department of Genetics and Genomic Sciences, Icahn Institute for Data Science and Genomic Technology, Icahn School of Medicine at Mount Sinai, New York, NY"
  - id            : "2"
    institution   : "Cancer Target Discovery and Development Network (Oregon Health and Science University, Portland, OR), National Cancer Institute, National Institutes of Health"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["manuscript.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
biblio-style      : "apa"
output            :
  papaja::apa6_pdf:
    citation_package: biblatex
---

```{r setup, include = FALSE}
library("papaja")
# install.packages("tinytex")
# tinytex::install_tinytex()
# install.packages("citr")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

[@aksoy2017]

# Introduction

# Results

# Discussion

\newpage

# Methods

We used `r cite_r("manuscript_r_references.bib")` for all our analyses.

\newpage

# References

```{r create_r-references}
r_refs(file = "manuscript_r_references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

Finally, the APA guidelines require a note at the start of the reference section that explains what an asterisk means. This note can be added at the end of the document as follows.

<div id = "refs"></div>
\endgroup
