---
title: "Manuscript"
author: "Sally Claridge"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: cerulean
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
    code_folding: hide
    pandoc_args: [
      "+RTS", "-K2048m",
      "-RTS"
    ]
---

Analysis of the Broad's Avana CRISPR (Broad Institute Cancer Dependency Map 2018, Meyers et al. 2017) and the Broad and Dana-Farber Cancer Institite's Achilles shRNA (MacFarland et al. 2018, Data Science 2018).

The Avana screen produced results using [CERES](https://depmap.org/ceres/) (Meyers et al. 2017) ([GitHub](https://github.com/cancerdatasci/ceres)), which generates gene dependency scores from sgRNA depletion scores from gene essentiality screens and eliminates bias arising from the effect of copy number variation on Cas9 DNA cleavage. The lower the CERES score, the higher the likelihood that the gene is essential in the associated cell line. Scores are scaled per cell line such that a score of 0 is the median effect of nonessential genes and -1 is the median effect of common core essential genes.

All annotation files (copy number, mutation status, and gene expression) (Consortium and Consortium 2015, Barretina et al. 2012) were downloaded from the [DepMap Data Portal](https://depmap.org/portal/download/).

# Set up

--------------------------------------------------------------------------------

```{r setup, include=FALSE}
set.seed(1234)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# num_cores <- parallel::detectCores()
options(mc.cores = parallel::detectCores(), expressions = 5e5)
```

```{r}
library(bsselectR)
library(data.table)
library(ggpubr)
library(ggsignif)
library(kableExtra)
library(reshape2)
library(tidyverse)
```

# Functions {.tabset .tabset-fade .tabset-pills}

--------------------------------------------------------------------------------

## Significance thresholding

```{r}
adj_signif <- function(df) {
  # Takes df from compare_means and creates a signif code column for adjusted p-vals
  df$p.signif <- ifelse(df$p.signif == "ns", NA, df$p.signif)
  df$p.signif.adj <- ifelse(df$p.adj <= 0.0001, "****",
                     ifelse(df$p.adj <= 0.001, "***",
                     ifelse(df$p.adj <= 0.01, "**",
                     ifelse(df$p.adj <= 0.05, "*", NA))))
  df$p.short <- formatC(df$p, format = "g", digits =  2)
  df$p.adj.short <- formatC(df$p.adj, format = "g", digits =  2)
  
  return(df)
}
```

## Wilcoxon tests by drug

```{r}
WilcoxonByDrug <- function(drug, dataset, data_name) {
  use_data <- filter(dataset, Drug == drug)
  sig <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Hugo_Symbol", data = use_data, method = "wilcox.test", p.adjust.method = "BH")
  sig <- adj_signif(sig)
  sig <- sig[order(sig$p),]
  sig$Drug <- drug
  sig$Dataset <- toupper(data_name)
  return(sig)
}


WilcoxonByDrugAllGenes <- function(melt_file, grid_file, data_name) {
  druglist <- as.character(unique(melt_file$Drug))
  sig_df <- NULL
  for(i in 1:length(druglist)) {
    print(paste0("Current drug in process: ", druglist[i], " (drug ", i, " of ", length(druglist), ")."))
    flush.console()
    melted <- filter(melt_file, Drug == druglist[i])
    use_data <- merge(grid_file, melted, by = "accession_id", all.x = TRUE)
    use_data <- merge(use_data, ccl_converter, by = "accession_id", all.x = TRUE)
    use_data <- merge(use_data, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
    use_data$Mutation_Status_Nonsilent <- ifelse(is.na(use_data$Mutation_Status_Nonsilent), "Other", use_data$Mutation_Status_Nonsilent)
    use_data$Hugo_Symbol <- factor(use_data$Hugo_Symbol)
    use_data$AUC <- as.numeric(use_data$AUC)
    use_data <- filter(use_data, !is.na(use_data$AUC))
    use_data$Drug_Gene <- paste0(use_data$Drug, "_", use_data$Hugo_Symbol)
    use_data_ptmuts <- filter(use_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
    use_data_ptmuts$Keep_PtMut_Tests <- ifelse(use_data_ptmuts$Reference_Allele_Length == 1 | is.na(use_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(use_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
    use_data_ptmuts <- filter(use_data_ptmuts, Keep_PtMut_Tests == TRUE)
    
    sig <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Hugo_Symbol", data = use_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
    sig <- adj_signif(sig)
    sig <- sig[order(sig$p),]
    sig <- filter(sig, p <= 0.05)
    sig$Drug <- druglist[i]
    sig$Dataset <- toupper(data_name)
    sig_df <- rbind(sig_df, sig)
  }
  return(sig_df)
}
```

## Make per-drug boxplots
```{r}
makeDrugBoxplots <- function(drug, dataset) {
  if(dataset == "CCLE" | dataset == "ccle") {
    use_data <- filter(ccle_data_g2p_genesfilt, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- ccle_signif_g2p_ByDrug[[drug]]
    label_text <- data.frame(p.signif = sig$p.signif, p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = g2p_genes)
  }
  else if(dataset == "CTRP" | dataset == "ctrp") {
    use_data <- filter(ctrp_data_g2p_genesfilt, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- ctrp_signif_g2p_ByDrug[[drug]]
    label_text <- data.frame(p.signif = sig$p.signif, p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = g2p_genes)
  }
  else if(dataset == "GDSC" | dataset == "gdsc") {
    use_data <- filter(gdsc_data_g2p_genesfilt, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- gdsc_signif_g2p_ByDrug[[drug]]
    label_text <- data.frame(p.signif = sig$p.signif, p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = g2p_genes)
  }
  else { return("Error: Invalid dataset.") }
    
  plot <- ggplot(data = use_data, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
    facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 1) +
    geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85), outlier.shape = 3, outlier.size = 0.5) +
    scale_fill_manual(values = use_color) +
    guides(color = FALSE) +
    geom_text(data = label_text, mapping = aes(x = 1.5, y = max(use_data$AUC), label = p.signif), nudge_y = 0.2) +
    theme(strip.text.x = element_text(size = 10, angle = 45), legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
    labs(fill = "Mutation Status", y = "AUC", x = "Mutation Status", title = paste0(toupper(dataset), ": ", drug))
  return(plot)
}
```


# Data management {.tabset .tabset-fade .tabset-pills}

--------------------------------------------------------------------------------

## D. Charytonowicz cell line converter

This comprehensive cancer cell line information curated by Daniel Charytonowicz.
```{r}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
```

## DepMap cell line metadata

```{r}
ccl_info <- read.delim("./data_munging/DepMap-2018q3-celllines.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$Primary.Disease <- gsub("\\\\", "", ccl_info$Primary.Disease)
ccl_info$Primary.Disease <- gsub("Ewings", "Ewing's", ccl_info$Primary.Disease)

# From figshare
crispr_meta <- read.delim("./data_munging/sample_info_18Q3_crispr.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[7] <- "CCLE_Name"
```

## Genomic annotations

### CCLE MAF file

For mutation calling, get paired gene name and cell line fields in a data frame. For this analysis, we don't care about how many mutations there are per gene or what type of mutations there are, so I didn't save more information. I took unique gene-cell line combinations since we only cared about mutation presence/absence. Add a `Mutation_Status` column denoting all entries in MAF files as mutations present in the associated cell lines.
```{r}
maf_raw <- read.delim("./data_munging/CCLE_DepMap_18q3_maf_20180718.txt.gz", header = TRUE, sep = "\t")
# Filter for cell lines in CRISPR screen
maf_raw <- filter(maf_raw, Broad_ID %in% unique(crispr_meta$Broad_ID))
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"

# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))

# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Other", "Mutant")
maf_df <- unique(subset(maf_df, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Mutation_Status_Nonsilent", "Reference_Allele_Length", "Tumor_Seq_Allele1_Length")))
```

### CCLE copy number file

```{r, eval=FALSE}
cn <- read.delim("./data_munging/public_18Q3_gene_cn_v2.csv.gz", sep = ",", check.names = FALSE, header = TRUE)

# Convert log2 ratios (log2(CN/2)) to CN
cn[2:ncol(cn)] <- lapply(cn[2:ncol(cn)], function(x) 2 * (2 ^ x))

# Remove Entrez gene IDs from colnames
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"

# Melt
cn_melt <- melt(data = cn, id.vars = "Broad_ID", measure.vars = colnames(cn[2:ncol(cn)]), variable.name = "Hugo_Symbol", value.name = "Copy_Number")

saveRDS(cn_melt, "./data_munging/rds/cn_melt_18Q3.rds", compress = "xz")
```

```{r, eval=FALSE}
cn_melt <- readRDS("./data_munging/rds/cn_melt_18Q3.rds")
```

### CCLE gene expression file

Gene expression data (Reads Per Kilobase of transcript, per Million mapped reads, RPKM).
```{r, eval=FALSE}
ge <- read.delim("./../crispr_lineages_giant_files/CCLE_DepMap_18q3_RNAseq_RPKM_20180718.gct.gz", skip = 2, header = TRUE, sep = "\t", check.names = FALSE)

# Edit columns
ge$Name <- NULL
colnames(ge)[1] <- "Hugo_Symbol"

# Melt
ge_melt <- melt(data = ge, id.vars = "Hugo_Symbol", measure.vars = colnames(ge[2:ncol(ge)]), value.name = "RPKM")
## Split variable column
ge_melt <- with(ge_melt, cbind(Hugo_Symbol, colsplit(variable, pattern = " ", names = c("CCLE_Name", "Broad_ID")), RPKM))
## Remove parentheses around Broad IDs
ge_melt$Broad_ID <- gsub("\\(|\\)", "", ge_melt$Broad_ID)

saveRDS(ge_melt, "./../crispr_lineages_giant_files/ge_melt_18Q3.rds", compress = "xz")
```

```{r, eval=FALSE}
ge_melt <- readRDS("./../crispr_lineages_giant_files/ge_melt_18Q3.rds")
```

## G2P

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_v2_10_10_2018.tsv", sep = "\t", header = TRUE)
g2p$Drug <- tolower(g2p$Drug)
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p <- subset(g2p, grepl("^CID", g2p$DrugID))

g2p_drug_genes <- as.character(unique(g2p$Drug_Gene[!is.na(g2p$Drug_Gene)]))
g2p_genes <- unique(sapply(strsplit(g2p_drug_genes, "_", fixed = TRUE), function(x) x[2]))
g2p_drugs <- unique(sapply(strsplit(g2p_drug_genes, "_", fixed = TRUE), function(x) x[1]))

g2p_genes
g2p_drugs
# Order and genes taken from pink heatmapo from Daniel's analyses
g2p_genes <- c("EGFR", "KRAS", "ERBB2", "ERBB3", "MET", "ALK", "ROS1", "PDGFRA", "ABL1", "KIT", "PDGFRB", "PDGFB", "BRAF", "G6PD", "RET", "BRCA1", "BRCA2", "UGT1A1", "TSC1", "TSC2", "AKT1", "IDH2", "DPYD", "ESR1", "CYP19A1", "TPMT", "FLT3", "NPM1", "DNMT3A", "PML", "PTCH1", "JAK2", "MYD88", "MGMT")
```

## CRISPR

### DepMap dependency probabilities

```{r, eval=FALSE}
dep <- read.delim("./data_munging/gene_dependency_18Q3.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(dep) <- gsub(" .*", "", colnames(dep))
```

### Merge CRISPR data and annotations

The latest CRISPR CERES score data (18Q3, August 2018) was pulled from the [DepMap Data Portal](https://depmap.org/portal/download/) (Broad Institute Cancer Dependency Map 2018, Meyers et al. 2017).
```{r, eval=TRUE}
crispr <- read.delim("./data_munging/gene_effect_18Q3.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr) <- gsub(" .*", "", colnames(crispr))
```

Merge annotation data:
```{r, eval=FALSE}
# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Hugo_Symbol", value.name = "Score")
# Melt dependency probabilities dataset for merging
dep_melt <- melt(dep, id.vars = "Broad_ID", measure.vars = colnames(dep)[2:ncol(dep)], variable.name = "Hugo_Symbol", value.name = "Dep_Prob")
# Merge dependency probabilities
crispr_melt <- merge(crispr_melt, dep_melt, by = c("Broad_ID", "Hugo_Symbol"), all.x = TRUE)
# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_info, by = "Broad_ID", all.x = TRUE)
crispr_melt <- merge(crispr_melt, crispr_meta, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
# Merge mutation annotations
crispr_muts <- merge(crispr_melt, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Other", Mutation_Status_Nonsilent))
# Summarize number of mutant and Other cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Nonsilent_Other = sum(Mutation_Status_Nonsilent == "Other"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))
# Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Hugo_Symbol")
# Add Color columns
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Other", "thistle", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)
# Cell line lineages
crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)
# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Hugo_Symbol", "Broad_ID"), all.x = TRUE)
# Gene expression (RPKM)
ge_filt <- filter(ge_melt, Hugo_Symbol %in% unique(crispr_data$Hugo_Symbol) & Broad_ID %in% unique(crispr_data$Broad_ID))
crispr_data <- merge(crispr_data, ge_filt, by = c("Hugo_Symbol", "Broad_ID", "CCLE_Name"), all.x = TRUE)
crispr_data$RPKM_log2 <- log2(crispr_data$RPKM + 0.0001)
saveRDS(crispr_data, "./../crispr_lineages_giant_files/crispr_data_18Q3.rds", compress = "xz")
```

```{r}
crispr_data <- readRDS("./../crispr_lineages_giant_files/crispr_data_18Q3.rds")

crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Hugo_Symbol %in% g2p_genes)
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
```

## Drug screens

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
maf_df_g2p <- filter(maf_df, Hugo_Symbol %in% g2p_genes)

# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ccle_data <- merge(ccle_grid, ccle_melt, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, ccl_converter, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
ccle_data$Mutation_Status_Nonsilent <- ifelse(is.na(ccle_data$Mutation_Status_Nonsilent), "Other", ccle_data$Mutation_Status_Nonsilent)
ccle_data$Hugo_Symbol <- factor(ccle_data$Hugo_Symbol)
ccle_data$AUC <- as.numeric(ccle_data$AUC)
ccle_data$Color_Nonsilent <- ifelse(ccle_data$Mutation_Status_Nonsilent == "Other", "palegreen3", "springgreen4")
ccle_data$Color_Nonsilent <- factor(ccle_data$Color_Nonsilent)
ccle_data <- filter(ccle_data, !is.na(ccle_data$AUC))
ccle_data$Drug_Gene <- paste0(ccle_data$Drug, "_", ccle_data$Hugo_Symbol)
saveRDS(ccle_data, "./data_munging/rds/ccle_data_18Q3_g2pgenes.rds", compress = "xz")

# CTRP
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.vars = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ctrp_data <- merge(ctrp_grid, ctrp_melt, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, ccl_converter, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
ctrp_data$Mutation_Status_Nonsilent <- ifelse(is.na(ctrp_data$Mutation_Status_Nonsilent), "Other", ctrp_data$Mutation_Status_Nonsilent)
ctrp_data$Hugo_Symbol <- factor(ctrp_data$Hugo_Symbol)
ctrp_data$AUC <- as.numeric(ctrp_data$AUC)
ctrp_data$Color_Nonsilent <- ifelse(ctrp_data$Mutation_Status_Nonsilent == "Other", "slategray3", "steelblue4")
ctrp_data$Color_Nonsilent <- factor(ctrp_data$Color_Nonsilent)
ctrp_data <- filter(ctrp_data, !is.na(ctrp_data$AUC))
ctrp_data$Drug_Gene <- paste0(ctrp_data$Drug, "_", ctrp_data$Hugo_Symbol)
saveRDS(ctrp_data, "./data_munging/rds/ctrp_data_18Q3_g2pgenes.rds", compress = "xz")

# GDSC
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC")
gdsc_grid <- expand.grid("accession_id" = unique(gdsc_melt$accession_id), "Hugo_Symbol" = g2p_genes)
gdsc_data <- merge(gdsc_grid, gdsc_melt, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, ccl_converter, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
gdsc_data$Mutation_Status_Nonsilent <- ifelse(is.na(gdsc_data$Mutation_Status_Nonsilent), "Other", gdsc_data$Mutation_Status_Nonsilent)
gdsc_data$Hugo_Symbol <- factor(gdsc_data$Hugo_Symbol)
gdsc_data$AUC <- as.numeric(gdsc_data$AUC)
gdsc_data$Color_Nonsilent <- ifelse(gdsc_data$Mutation_Status_Nonsilent == "Other", "paleturquoise3", "turquoise4")
gdsc_data$Color_Nonsilent <- factor(gdsc_data$Color_Nonsilent)
gdsc_data <- filter(gdsc_data, !is.na(gdsc_data$AUC))
gdsc_data$Drug_Gene <- paste0(gdsc_data$Drug, "_", gdsc_data$Hugo_Symbol)
saveRDS(gdsc_data, "./data_munging/rds/gdsc_data_18Q3_g2pgenes.rds", compress = "xz")
```

```{r}
ccle_data <- readRDS("./data_munging/rds/ccle_data_18Q3_g2pgenes.rds")
ccle_data_ptmuts <- filter(ccle_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ccle_data_ptmuts$Keep_PtMut_Tests <- ifelse(ccle_data_ptmuts$Reference_Allele_Length == 1 | is.na(ccle_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ccle_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ccle_data_ptmuts <- filter(ccle_data_ptmuts, Keep_PtMut_Tests == TRUE)

ctrp_data <- readRDS("./data_munging/rds/ctrp_data_18Q3_g2pgenes.rds")
ctrp_data_ptmuts <- filter(ctrp_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ctrp_data_ptmuts$Keep_PtMut_Tests <- ifelse(ctrp_data_ptmuts$Reference_Allele_Length == 1 | is.na(ctrp_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ctrp_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ctrp_data_ptmuts <- filter(ctrp_data_ptmuts, Keep_PtMut_Tests == TRUE)

gdsc_data <- readRDS("./data_munging/rds/gdsc_data_18Q3_g2pgenes.rds")
gdsc_data_ptmuts <- filter(gdsc_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
gdsc_data_ptmuts$Keep_PtMut_Tests <- ifelse(gdsc_data_ptmuts$Reference_Allele_Length == 1 | is.na(gdsc_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(gdsc_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
gdsc_data_ptmuts <- filter(gdsc_data_ptmuts, Keep_PtMut_Tests == TRUE)

# Filter drug screen datasets for G2P drug-gene associations
ccle_data_g2p <- filter(ccle_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
ctrp_data_g2p <- filter(ctrp_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
gdsc_data_g2p <- filter(gdsc_data_ptmuts, Drug_Gene %in% g2p_drug_genes)

# Filter drug screen datasets for G2P genes in drug-gene associations
ccle_data_g2p_genesfilt <- filter(ccle_data_ptmuts, Hugo_Symbol %in% g2p_genes)
ccle_data_g2p_genesfilt$Hugo_Symbol <- ordered(ccle_data_g2p_genesfilt$Hugo_Symbol, levels = g2p_genes)
ctrp_data_g2p_genesfilt <- filter(ctrp_data_ptmuts, Hugo_Symbol %in% g2p_genes)
ctrp_data_g2p_genesfilt$Hugo_Symbol <- ordered(ctrp_data_g2p_genesfilt$Hugo_Symbol, levels = g2p_genes)
gdsc_data_g2p_genesfilt <- filter(gdsc_data_ptmuts, Hugo_Symbol %in% g2p_genes)
gdsc_data_g2p_genesfilt$Hugo_Symbol <- ordered(gdsc_data_g2p_genesfilt$Hugo_Symbol, levels = g2p_genes)
```

# Wilcoxon tests

--------------------------------------------------------------------------------

## CRISPR: G2P genes

```{r, eval=FALSE}
crispr_signif_g2p_gene <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol"), data = crispr_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_g2p_gene <- adj_signif(crispr_signif_g2p_gene)
crispr_signif_g2p_gene <- crispr_signif_g2p_gene[order(crispr_signif_g2p_gene$p),]
saveRDS(crispr_signif_g2p_gene, "./data_munging/rds/crispr_signif_g2p_gene.rds")
```

```{r, eval=TRUE}
crispr_signif_g2p_gene <- readRDS("./data_munging/rds/crispr_signif_g2p_gene.rds")

crispr_signif_g2p_20gene <- as.character(crispr_signif_g2p_gene$Hugo_Symbol[0:20])

knitr::kable(crispr_signif_g2p_gene[, c("Hugo_Symbol", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results for Level A point mutations (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=7, fig.width=12}
crispr_data_signif_g2p_gene_filt <- filter(crispr_data_ptmuts, Hugo_Symbol %in% crispr_signif_g2p_20gene)
crispr_data_signif_g2p_gene <- merge(crispr_data_signif_g2p_gene_filt, crispr_signif_g2p_gene, by = "Hugo_Symbol")
crispr_data_signif_g2p_gene$Hugo_Symbol <- ordered(crispr_data_signif_g2p_gene$Hugo_Symbol, levels = crispr_signif_g2p_20gene)
crispr_data_signif_g2p_gene_color <- as.character(crispr_data_signif_g2p_gene$Color_Nonsilent)
names(crispr_data_signif_g2p_gene_color) <- crispr_data_signif_g2p_gene$Mutation_Status_Nonsilent
crispr_data_signif_g2p_gene <- merge(crispr_data_signif_g2p_gene, crispr_data_signif_g2p_gene %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")

crispr_data_signif_g2p_gene_text <- unique(data.frame(p.signif = crispr_data_signif_g2p_gene$p.signif, p.signif.adj = crispr_data_signif_g2p_gene$p.signif.adj,  Hugo_Symbol = crispr_data_signif_g2p_gene$Hugo_Symbol, Text_y = crispr_data_signif_g2p_gene$Text_y))

crispr_data_signif_g2p_gene_plot <- ggplot(data = crispr_data_signif_g2p_gene, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 2, scales = "free_y") +
  geom_jitter(aes(color = Mutation_Status_Nonsilent), alpha = 0.5, width = 0.3) +
  geom_boxplot(color = "black", fill = NA, outlier.shape = NA) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = crispr_data_signif_g2p_gene_color) + 
  geom_text(data = crispr_data_signif_g2p_gene_text, mapping = aes(x = 1.5, y = Text_y, label = p.signif), nudge_y = 0.2) +
  theme(legend.position = "top", axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank()) +
  labs(y = "CERES Score",
       color = "Mutation Status",
       title = "CERES scores for select Level A G2P genes",
       subtitle = "Genes are sorted by increasing p-value.")
crispr_data_signif_g2p_gene_plot

# ggsave("./plots_18Q3/manuscript/crispr_data_signif_g2p_gene_plot.png", crispr_data_signif_g2p_gene_plot, width = 12, height = 7, units = "in")
```

## CRISPR: G2P genes and lineage

```{r, eval=FALSE}
crispr_signif_g2p_lineage <- compare_means(Score ~ Mutation_Status_Nonsilent, group.by = c("Hugo_Symbol", "group_general_lineage_name"), data = crispr_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
crispr_signif_g2p_lineage <- adj_signif(crispr_signif_g2p_lineage)
crispr_signif_g2p_lineage <- crispr_signif_g2p_lineage[order(crispr_signif_g2p_lineage$p),]
saveRDS(crispr_signif_g2p_lineage, "./data_munging/rds/crispr_signif_g2p_lineage.rds")

# write.table(crispr_signif_g2p_lineage, file = "~/Desktop/crispr_signif_g2p_lineage.csv", quote = FALSE, sep = ",", row.names = FALSE)
```

```{r, eval=TRUE}
crispr_signif_g2p_lineage <- readRDS("./data_munging/rds/crispr_signif_g2p_lineage.rds")

knitr::kable(filter(crispr_signif_g2p_lineage, p < 0.1)[, c("Hugo_Symbol", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "Wilcoxon test results comparing non-silent mutant vs other cell lines by lineage, p < 0.1 (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r fig.height=15, fig.width=15}
crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_gene_filt, crispr_signif_g2p_lineage, by = c("Hugo_Symbol", "group_general_lineage_name"))
crispr_data_signif_g2p_lineage$Hugo_Symbol <- ordered(crispr_data_signif_g2p_lineage$Hugo_Symbol, levels = crispr_signif_g2p_20gene)
crispr_data_signif_g2p_lineage_color <- as.character(crispr_data_signif_g2p_lineage$Color_Nonsilent)
names(crispr_data_signif_g2p_lineage_color) <- crispr_data_signif_g2p_lineage$Mutation_Status_Nonsilent
crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_lineage, crispr_data_signif_g2p_lineage %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")

crispr_data_signif_g2p_lineage_text <- unique(data.frame(p.signif = crispr_data_signif_g2p_lineage$p.signif, p.signif.adj = crispr_data_signif_g2p_lineage$p.signif.adj,  group_general_lineage_name = crispr_data_signif_g2p_lineage$group_general_lineage_name, Hugo_Symbol = crispr_data_signif_g2p_lineage$Hugo_Symbol, Text_y = crispr_data_signif_g2p_lineage$Text_y))

crispr_data_signif_g2p_lineage_plot <- ggplot(data = crispr_data_signif_g2p_lineage, aes(x = group_general_lineage_name, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 4, scales = "free_y") +
  geom_point(aes(color = Mutation_Status_Nonsilent), alpha = 0.5) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = crispr_data_signif_g2p_lineage_color) +
  geom_text(data = crispr_data_signif_g2p_lineage_text, mapping = aes(x = group_general_lineage_name, y = Text_y, label = p.signif), angle = 90, vjust = 0.8, nudge_y = 0.1) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES score by cell line lineage for significant Level A G2P genes",
       subtitle = "* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing lineage-agnostic p-value.")
crispr_data_signif_g2p_lineage_plot

# ggsave("./plots_18Q3/manuscript/crispr_crispr_data_signif_g2p_lineage_plot.png", crispr_data_signif_g2p_lineage_plot, width = 12, height = 15, units = "in")
```

```{r}
crispr_data_signif_g2p_lineage_summ <- crispr_data_signif_g2p_lineage %>% group_by(Hugo_Symbol, group_general_lineage_name, Mutation_Status_Nonsilent) %>% tally()
# write.table(crispr_data_signif_g2p_lineage_summ, file = "~/Desktop/crispr_data_signif_g2p_lineage_summ_20181005.tsv", quote = FALSE, sep = "\t")

knitr::kable(crispr_data_signif_g2p_lineage_summ, caption = "Mutant/Other counts for gene-lineage combinations") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## CCLE: G2P gene and drug

```{r, eval=FALSE}
ccle_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ccle_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ccle_signif_g2p <- adj_signif(ccle_signif_g2p)
ccle_signif_g2p <- ccle_signif_g2p[order(ccle_signif_g2p$p),]
saveRDS(ccle_signif_g2p, "./data_munging/rds/ccle_signif_g2p_gene.rds")

ccle_signif_g2p_ByDrug <- lapply(unique(ccle_data_g2p_genesfilt$Drug), WilcoxonByDrug, dataset = ccle_data_g2p_genesfilt, data_name = "ccle")
names(ccle_signif_g2p_ByDrug) <- unique(ccle_data_g2p_genesfilt$Drug)
saveRDS(ccle_signif_g2p_ByDrug, "./data_munging/rds/ccle_signif_g2p_ByDrug.rds")
```

```{r, eval=TRUE}
ccle_signif_g2p <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")

knitr::kable(ccle_signif_g2p[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for level A G2P drug-gene associations (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r}
ccle_signif_g2p_ByDrug <- readRDS("./data_munging/rds/ccle_signif_g2p_ByDrug.rds")
ccle_signif_g2p_ByDrug_all <- rbindlist(ccle_signif_g2p_ByDrug, use.names = TRUE)
ccle_signif_g2p_ByDrug_all <- ccle_signif_g2p_ByDrug_all[order(ccle_signif_g2p_ByDrug_all$p),]

knitr::kable(ccle_signif_g2p_ByDrug_all[, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: By-drug Wilcoxon test results for point mutations in level A G2P genes (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## CTRP: G2P gene and drug

```{r, eval=FALSE}
ctrp_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ctrp_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ctrp_signif_g2p <- adj_signif(ctrp_signif_g2p)
ctrp_signif_g2p <- ctrp_signif_g2p[order(ctrp_signif_g2p$p),]
saveRDS(ctrp_signif_g2p, "./data_munging/rds/ctrp_signif_g2p_gene.rds")

ctrp_signif_g2p_ByDrug <- lapply(unique(ctrp_data_g2p_genesfilt$Drug), WilcoxonByDrug, dataset = ctrp_data_g2p_genesfilt, data_name = "CTRP")
names(ctrp_signif_g2p_ByDrug) <- unique(ctrp_data_g2p_genesfilt$Drug)
saveRDS(ctrp_signif_g2p_ByDrug, "./data_munging/rds/ctrp_signif_g2p_ByDrug.rds")
```

```{r}
ctrp_signif_g2p <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")

knitr::kable(ctrp_signif_g2p[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: Wilcoxon test results for level A G2P drug-gene associations (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r}
ctrp_signif_g2p_ByDrug <- readRDS("./data_munging/rds/ctrp_signif_g2p_ByDrug.rds")
ctrp_signif_g2p_ByDrug_all <- rbindlist(ctrp_signif_g2p_ByDrug, use.names = TRUE)
ctrp_signif_g2p_ByDrug_all <- ctrp_signif_g2p_ByDrug_all[order(ctrp_signif_g2p_ByDrug_all$p),]

knitr::kable(ctrp_signif_g2p_ByDrug_all[, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: By-drug Wilcoxon test results for point mutations in level A G2P genes (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## GDSC: G2P gene and drug

```{r, eval=FALSE}
gdsc_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = gdsc_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
gdsc_signif_g2p <- adj_signif(gdsc_signif_g2p)
gdsc_signif_g2p <- gdsc_signif_g2p[order(gdsc_signif_g2p$p),]
saveRDS(gdsc_signif_g2p, "./data_munging/rds/gdsc_signif_g2p_gene.rds")

gdsc_signif_g2p_ByDrug <- lapply(unique(gdsc_data_g2p_genesfilt$Drug), WilcoxonByDrug, dataset = gdsc_data_g2p_genesfilt, data_name = "gdsc")
names(gdsc_signif_g2p_ByDrug) <- unique(gdsc_data_g2p_genesfilt$Drug)
saveRDS(gdsc_signif_g2p_ByDrug, "./data_munging/rds/gdsc_signif_g2p_ByDrug.rds")
```

```{r}
gdsc_signif_g2p <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")

knitr::kable(gdsc_signif_g2p[, c("Drug_Gene", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: Wilcoxon test results for level A G2P drug-gene associations (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

```{r}
gdsc_signif_g2p_ByDrug <- readRDS("./data_munging/rds/gdsc_signif_g2p_ByDrug.rds")
gdsc_signif_g2p_ByDrug_all <- rbindlist(gdsc_signif_g2p_ByDrug, use.names = TRUE)
gdsc_signif_g2p_ByDrug_all <- gdsc_signif_g2p_ByDrug_all[order(gdsc_signif_g2p_ByDrug_all$p),]

knitr::kable(gdsc_signif_g2p_ByDrug_all[, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: by-drug Wilcoxon test results for point mutations in level A G2P genes (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## Drug screens: All CRISPR genes

Drug metadata:
```{r}
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
```

CCLE:

user   system  elapsed 
1483.399  262.649 1901.698 

```{r, eval=FALSE}
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "AUC")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
system.time({ ccle_data_allgenes <- WilcoxonByDrugAllGenes(ccle_melt, ccle_grid, "CCLE") })
saveRDS(ccle_data_allgenes, "./data_munging/rds/ccle_data_18Q3_allgenes.rds", compress = "xz")
```

```{r}
ccle_data_allgenes <- readRDS("./data_munging/rds/ccle_data_18Q3_allgenes.rds")
ccle_data_allgenes <- ccle_data_allgenes[order(ccle_data_allgenes$p),]

knitr::kable(ccle_data_allgenes[1:1000, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: by-drug Wilcoxon test results for point mutations in genes targeted in CRISPR screen (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

CTRP:
```{r, eval=FALSE}
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.var  s = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
ctrp_data_allgenes <- WilcoxonByDrugAllGenes(ctrp_melt, ctrp_grid, "CTRP")
saveRDS(ctrp_data_allgenes, "./data_munging/rds/ctrp_data_18Q3_allgenes.rds", compress = "xz")
```

```{r}
ctrp_data_allgenes <- readRDS("./data_munging/rds/ctrp_data_18Q3_allgenes.rds")
ctrp_data_allgenes <- ctrp_data_allgenes[order(ctrp_data_allgenes$p),]

knitr::kable(ctrp_data_allgenes[1:1000, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: by-drug Wilcoxon test results for point mutations in genes targeted in CRISPR screen (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

GDSC:
```{r, eval=FALSE}
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC")
gdsc_grid <- expand.grid("accession_id" = unique(gdsc_melt$accession_id), "Hugo_Symbol" = as.character(unique(crispr_data$Hugo_Symbol)))
system.time({ gdsc_data_allgenes <- WilcoxonByDrugAllGenes(gdsc_melt, gdsc_grid, "GDSC") })
saveRDS(gdsc_data_allgenes, "./data_munging/rds/gdsc_data_18Q3_allgenes.rds", compress = "xz")
```

```{r}
gdsc_data_allgenes <- readRDS("./data_munging/rds/gdsc_data_18Q3_allgenes.rds")
gdsc_data_allgenes <- gdsc_data_allgenes[order(gdsc_data_allgenes$p),]

knitr::kable(gdsc_data_allgenes[1:1000, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: by-drug Wilcoxon test results for point mutations in genes targeted in CRISPR screen (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

# Dataset summaries

--------------------------------------------------------------------------------

## CRISPR: Mutation status counts

```{r}
crispr_data_g2p_summ <- unique(crispr_data_g2p[, c("Hugo_Symbol", "N_Nonsilent_Mutant", "N_Nonsilent_Other")])
crispr_data_g2p_summ$Percent_Mutant <- crispr_data_g2p_summ$N_Nonsilent_Mutant / (crispr_data_g2p_summ$N_Nonsilent_Mutant + crispr_data_g2p_summ$N_Nonsilent_Other) * 100
rownames(crispr_data_g2p_summ) <- NULL
# write.table(crispr_data_g2p_summ, file = "~/Desktop/crispr_data_g2p_summ_20181003.tsv", quote = FALSE, sep = "\t")

knitr::kable(crispr_data_g2p_summ, caption = "CRISPR: Level A G2P genes mutation status summary") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

## Drugscreen G2P evidence

```{r fig.height=4, fig.width=8}
g2p_druggene_p_tally <- g2p %>% group_by(Drug_Gene) %>% tally()
g2p_druggene_p_summ <- merge(g2p_druggene_p_tally, ccle_signif_g2p[, c("Drug_Gene", "p")], by = "Drug_Gene", all.x = TRUE)
g2p_druggene_p_summ <- merge(g2p_druggene_p_summ, ctrp_signif_g2p[, c("Drug_Gene", "p")], by = "Drug_Gene", all.x = TRUE, suffixes = c("_CCLE", "_CTRP"))
g2p_druggene_p_summ <- merge(g2p_druggene_p_summ, gdsc_signif_g2p[, c("Drug_Gene", "p")], by = "Drug_Gene", all.x = TRUE)
colnames(g2p_druggene_p_summ)[colnames(g2p_druggene_p_summ) == "p_CCLE"] <- "CCLE"
colnames(g2p_druggene_p_summ)[colnames(g2p_druggene_p_summ) == "p_CTRP"] <- "CTRP"
colnames(g2p_druggene_p_summ)[colnames(g2p_druggene_p_summ) == "p"] <- "GDSC"
g2p_druggene_p_summ_melt <- melt(g2p_druggene_p_summ, id.vars = c("Drug_Gene", "n"), measure.vars = c("CCLE", "CTRP", "GDSC"), variable.name = "Dataset", value.name = "p")

g2p_druggene_p_summ_plot <- ggplot(data = g2p_druggene_p_summ_melt, aes(x = p, y = n, shape = Dataset, color = Dataset)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("green3", "darkslategray4", "darkturquoise")) +
  labs(x = "Gene-drug association p-value by mutation status", y = "Number of level A G2P associations")
g2p_druggene_p_summ_plot
# ggsave("./plots_18Q3/manuscript/g2p_druggene_p_summ_plot.png", g2p_druggene_p_summ_plot, device = "png", dpi = 450, width = 8, height = 4, units = "in")
```

```{r}
ccle_druggene_mad_summ <- ccle_data_ptmuts %>% group_by(Drug_Gene, Mutation_Status_Nonsilent) %>% summarize(Mean_AUC = mean(AUC))
ccle_druggene_mad_summ <- dcast(ccle_druggene_mad_summ, Drug_Gene ~ Mutation_Status_Nonsilent)
ccle_druggene_mad_summ$MAD <- abs(ccle_druggene_mad_summ$Mutant - ccle_druggene_mad_summ$Other)
ccle_druggene_p_summ <- ccle_signif_g2p_ByDrug_all
ccle_druggene_p_summ$Drug_Gene <- paste0(ccle_druggene_p_summ$Drug, "_", ccle_druggene_p_summ$Hugo_Symbol)
ccle_druggene_summ <- merge(merge(g2p_druggene_p_tally, ccle_druggene_p_summ[, c("Drug_Gene", "p")], by = "Drug_Gene", all.y = TRUE), ccle_druggene_mad_summ, by = "Drug_Gene", all = TRUE)
ccle_druggene_summ$n <- ifelse(is.na(ccle_druggene_summ$n), 0, ccle_druggene_summ$n)
ccle_druggene_summ$InG2P <- ifelse(ccle_druggene_summ$n != 0, "Yes", "No")
ccle_druggene_summ$Dataset <- "CCLE"
ccle_p_ttest <- t.test(ccle_druggene_summ$p ~ ccle_druggene_summ$InG2P)
ccle_p_ttest_res <- paste0("t = ", round(ccle_p_ttest$statistic, 3), ", df = ", round(ccle_p_ttest$parameter, 3), ", p = ", round(ccle_p_ttest$p.value, 3))
ccle_mad_ttest <- t.test(ccle_druggene_summ$MAD ~ ccle_druggene_summ$InG2P)
ccle_mad_ttest_res <- paste0("t = ", round(ccle_mad_ttest$statistic, 3), ", df = ", round(ccle_mad_ttest$parameter, 3), ", p = ", round(ccle_mad_ttest$p.value, 3))

ctrp_druggene_mad_summ <- ctrp_data_ptmuts %>% group_by(Drug_Gene, Mutation_Status_Nonsilent) %>% summarize(Mean_AUC = mean(AUC))
ctrp_druggene_mad_summ <- dcast(ctrp_druggene_mad_summ, Drug_Gene ~ Mutation_Status_Nonsilent)
ctrp_druggene_mad_summ$MAD <- abs(ctrp_druggene_mad_summ$Mutant - ctrp_druggene_mad_summ$Other)
ctrp_druggene_p_summ <- ctrp_signif_g2p_ByDrug_all
ctrp_druggene_p_summ$Drug_Gene <- paste0(ctrp_druggene_p_summ$Drug, "_", ctrp_druggene_p_summ$Hugo_Symbol)
ctrp_druggene_summ <- merge(merge(g2p_druggene_p_tally, ctrp_druggene_p_summ[, c("Drug_Gene", "p")], by = "Drug_Gene", all.y = TRUE), ctrp_druggene_mad_summ, by = "Drug_Gene", all = TRUE)
ctrp_druggene_summ$n <- ifelse(is.na(ctrp_druggene_summ$n), 0, ctrp_druggene_summ$n)
ctrp_druggene_summ$InG2P <- ifelse(ctrp_druggene_summ$n != 0, "Yes", "No")
ctrp_druggene_summ$Dataset <- "CTRP"
ctrp_p_ttest <- t.test(ctrp_druggene_summ$p ~ ctrp_druggene_summ$InG2P)
ctrp_p_ttest_res <- paste0("t = ", round(ctrp_p_ttest$statistic, 3), ", df = ", round(ctrp_p_ttest$parameter, 3), ", p = ", round(ctrp_p_ttest$p.value, 3))
ctrp_mad_ttest <- t.test(ctrp_druggene_summ$MAD ~ ctrp_druggene_summ$InG2P)
ctrp_mad_ttest_res <- paste0("t = ", round(ctrp_mad_ttest$statistic, 3), ", df = ", round(ctrp_mad_ttest$parameter, 3), ", p = ", round(ctrp_mad_ttest$p.value, 3))

gdsc_druggene_mad_summ <- gdsc_data_ptmuts %>% group_by(Drug_Gene, Mutation_Status_Nonsilent) %>% summarize(Mean_AUC = mean(AUC))
gdsc_druggene_mad_summ <- dcast(gdsc_druggene_mad_summ, Drug_Gene ~ Mutation_Status_Nonsilent)
gdsc_druggene_mad_summ$MAD <- abs(gdsc_druggene_mad_summ$Mutant - gdsc_druggene_mad_summ$Other)
gdsc_druggene_p_summ <- gdsc_signif_g2p_ByDrug_all
gdsc_druggene_p_summ$Drug_Gene <- paste0(gdsc_druggene_p_summ$Drug, "_", gdsc_druggene_p_summ$Hugo_Symbol)
gdsc_druggene_summ <- merge(merge(g2p_druggene_p_tally, gdsc_druggene_p_summ[, c("Drug_Gene", "p")], by = "Drug_Gene", all.y = TRUE), gdsc_druggene_mad_summ, by = "Drug_Gene", all = TRUE)
gdsc_druggene_summ$n <- ifelse(is.na(gdsc_druggene_summ$n), 0, gdsc_druggene_summ$n)
gdsc_druggene_summ$InG2P <- ifelse(gdsc_druggene_summ$n != 0, "Yes", "No")
gdsc_druggene_summ$Dataset <- "GDSC"
gdsc_p_ttest <- t.test(gdsc_druggene_summ$p ~ gdsc_druggene_summ$InG2P)
gdsc_p_ttest_res <- paste0("t = ", round(gdsc_p_ttest$statistic, 3), ", df = ", round(gdsc_p_ttest$parameter, 3), ", p = ", round(gdsc_p_ttest$p.value, 3))
gdsc_mad_ttest <- t.test(gdsc_druggene_summ$MAD ~ gdsc_druggene_summ$InG2P)
gdsc_mad_ttest_res <- paste0("t = ", round(gdsc_mad_ttest$statistic, 3), ", df = ", round(gdsc_mad_ttest$parameter, 3), ", p = ", round(gdsc_mad_ttest$p.value, 3))

g2p_druggene_all_summ <- rbind(ccle_druggene_summ, ctrp_druggene_summ, gdsc_druggene_summ)
g2p_druggene_all_summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"), 
                                         Test_Results_p = c(ccle_p_ttest_res, ctrp_p_ttest_res, gdsc_p_ttest_res),
                                         Test_Results_MAD = c(ccle_mad_ttest_res, ctrp_mad_ttest_res, gdsc_mad_ttest_res))
```

```{r fig.height=7, fig.width=12}
## MAD
plot_mad_point <- ggplot(data = g2p_druggene_all_summ) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_point(mapping = aes(color = InG2P, x = MAD, y = n), alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 110, by = 10), labels = seq(0, 110, by = 10)) +
  geom_text(data = g2p_druggene_all_summ_text, mapping = aes(x = c(0.75, 4, 0.19), y = 110, label = Test_Results_MAD)) +
  labs(x = "Gene-drug association difference in mean AUC between mutation groups", y = "Number of level A G2P associations", color = "Does the gene-drug combination have level A G2P indication?") +
  theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())

plot_mad_box <- ggplot(data = g2p_druggene_all_summ) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_boxplot(mapping = aes(color = InG2P, y = MAD, x = InG2P), alpha = 0.5) +
  coord_flip() +
  labs(x = "Level A G2P\nassociation") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

plot_mad_point <- ggplot_gtable(ggplot_build(plot_mad_point))
plot_mad_box <- ggplot_gtable(ggplot_build(plot_mad_box))
mad_maxWidth <-  grid::unit.pmax(plot_mad_point$widths, plot_mad_box$widths)
plot_mad_point$widths <- as.list(mad_maxWidth)
plot_mad_box$widths <- as.list(mad_maxWidth)

g2p_druggene_all_mad_summ_plot <- cowplot::plot_grid(plot_mad_box,
                   plot_mad_point,
                   nrow = 2,
                   rel_heights = c(1, 4))
g2p_druggene_all_mad_summ_plot
# ggsave("./plots_18Q3/manuscript/g2p_druggene_all_mad_summ_plot.png", g2p_druggene_all_mad_summ_plot, device = "png", dpi = 450, width = 12, height = 7, units = "in")

## P-value
plot_p_point <- ggplot(data = g2p_druggene_all_summ) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_point(mapping = aes(color = InG2P, x = p, y = n), alpha = 0.5) +
  scale_y_continuous(breaks = seq(0, 110, by = 10), labels = seq(0, 110, by = 10)) +
  geom_text(data = g2p_druggene_all_summ_text, mapping = aes(x = 0.5, y = 110, label = Test_Results_p)) +
  labs(x = "Gene-drug association p-value between mutation groups", y = "Number of level A G2P associations", color = "Does the gene-drug combination have level A G2P indication?") +
  theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())

plot_p_box <- ggplot(data = g2p_druggene_all_summ) +
  facet_wrap(~ Dataset, scales = "free_x") +
  geom_boxplot(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.5) +
  coord_flip() +
  labs(x = "Level A G2P\nassociation") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

plot_p_point <- ggplot_gtable(ggplot_build(plot_p_point))
plot_p_box <- ggplot_gtable(ggplot_build(plot_p_box))
p_maxWidth <-  grid::unit.pmax(plot_p_point$widths, plot_p_box$widths)
plot_p_point$widths <- as.list(p_maxWidth)
plot_p_box$widths <- as.list(p_maxWidth)

g2p_druggene_all_p_summ_plot <- cowplot::plot_grid(plot_p_box,
                   plot_p_point,
                   nrow = 2,
                   rel_heights = c(1, 4))
g2p_druggene_all_p_summ_plot
# ggsave("./plots_18Q3/manuscript/g2p_druggene_all_p_summ_plot.png", g2p_druggene_all_p_summ_plot, device = "png", dpi = 450, width = 12, height = 7, units = "in")
```

```{r}
crispr_ccl <- unique(select(crispr_data, Broad_ID, CCLE_Name))
ccle_ccl <- unique(select(ccle_data, Broad_ID, CCLE_Name, accession_id))
ctrp_ccl <- unique(select(ctrp_data, Broad_ID, CCLE_Name, accession_id))
gdsc_ccl <- unique(select(gdsc_data, Broad_ID, CCLE_Name, accession_id))

# all_ccl <- merge(ccle_ccl, ctrp_ccl, by = c("Broad_ID", "CCLE_Name", "accession_id"), all = TRUE)
```


# Boxplots

--------------------------------------------------------------------------------

Mutation status boxplots for genes identified in G2P.

## Figure 1

Boxplots ordered by gene based on frequency of G2P associations.

### CRISPR

```{r fig.height=5, fig.width=20}
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p_color <- as.character(crispr_data_g2p$Color_Nonsilent)
names(crispr_data_g2p_color) <- crispr_data_g2p$Mutation_Status_Nonsilent

crispr_label_text <- data.frame(p.signif.adj = crispr_signif_g2p_gene$p.signif.adj, p.signif = crispr_signif_g2p_gene$p.signif, Hugo_Symbol = crispr_signif_g2p_gene$Hugo_Symbol)
crispr_label_text <- filter(crispr_label_text, Hugo_Symbol %in% g2p_genes)
crispr_label_text$Hugo_Symbol <- factor(crispr_label_text$Hugo_Symbol, levels = g2p_genes)

crispr_data_g2p_plot <- ggplot(data = crispr_data_g2p, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 1) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  scale_fill_manual(values = crispr_data_g2p_color) +
  guides(color = FALSE) +
  geom_text(data = crispr_label_text, mapping = aes(x = 1.5, y = 2, label = p.signif), nudge_y = 0.1) +
  theme(strip.text.x = element_text(size = 10, angle = 45), legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "CRISPR", fill = "Mutation Status", y = "CERES Score", x = "Mutation Status")
crispr_data_g2p_plot

# ggsave("./plots_18Q3/manuscript/crispr_data_g2p_plot.png", crispr_data_g2p_plot, device = "png", dpi = 450, width = 20, height = 5, units = "in")
```

### CCLE

Plot code:
```{r, eval=FALSE}
ccle_drugs <- as.character(unique(ccle_data_g2p_genesfilt$Drug))
ccle_g2p_drugboxplots <- lapply(ccle_drugs, makeDrugBoxplots, dataset = "CCLE")
ccle_g2p_drugboxplots_paths <- paste0(ccle_drugs, "_ccle_drugboxplots.png")
pwalk(list(ccle_g2p_drugboxplots_paths, ccle_g2p_drugboxplots), ggsave, path = "./plots_18Q3/manuscript/ccle_g2p_drugboxplots", dpi = 450, width = 20, height = 5, units = "in")
```

Data management:
```{r, eval=TRUE}
ccle_g2p_drugboxplots <- paste0(list.files("./plots_18Q3/manuscript/ccle_g2p_drugboxplots", full.names = TRUE))
names(ccle_g2p_drugboxplots) <- str_replace_all(ccle_g2p_drugboxplots, c("_ccle_drugboxplots.png" = "", "./plots_18Q3/manuscript/ccle_g2p_drugboxplots/" = ""))
bsselect(ccle_g2p_drugboxplots, type = "img", live_search = TRUE, show_tick = TRUE, height = 100)
```

### CTRP

Plot code:
```{r, eval=FALSE}
ctrp_drugs <- as.character(unique(ctrp_data_g2p_genesfilt$Drug))
ctrp_g2p_drugboxplots <- lapply(ctrp_drugs, makeDrugBoxplots, dataset = "ctrp")
ctrp_g2p_drugboxplots_paths <- paste0(ctrp_drugs, "_ctrp_drugboxplots.png")
pwalk(list(ctrp_g2p_drugboxplots_paths, ctrp_g2p_drugboxplots), ggsave, path = "./plots_18Q3/manuscript/ctrp_g2p_drugboxplots", dpi = 450, width = 20, height = 5, units = "in")
```

Data management:
```{r, eval=TRUE}
ctrp_g2p_drugboxplots <- paste0(list.files("./plots_18Q3/manuscript/ctrp_g2p_drugboxplots", full.names = TRUE))
names(ctrp_g2p_drugboxplots) <- str_replace_all(ctrp_g2p_drugboxplots, c("_ctrp_drugboxplots.png" = "", "./plots_18Q3/manuscript/ctrp_g2p_drugboxplots/" = ""))
bsselect(ctrp_g2p_drugboxplots, type = "img", live_search = TRUE, show_tick = TRUE, height = 100)
```

### GDSC

Plot code:
```{r, eval=FALSE}
gdsc_drugs <- as.character(unique(gdsc_data_g2p_genesfilt$Drug))
gdsc_g2p_drugboxplots <- lapply(gdsc_drugs, makeDrugBoxplots, dataset = "gdsc")
gdsc_g2p_drugboxplots_paths <- paste0(gdsc_drugs, "_gdsc_drugboxplots.png")
pwalk(list(gdsc_g2p_drugboxplots_paths, gdsc_g2p_drugboxplots), ggsave, path = "./plots_18Q3/manuscript/gdsc_g2p_drugboxplots", dpi = 450, width = 20, height = 5, units = "in")
```

Data management:
```{r, eval=TRUE}
gdsc_g2p_drugboxplots <- paste0(list.files("./plots_18Q3/manuscript/gdsc_g2p_drugboxplots", full.names = TRUE))
names(gdsc_g2p_drugboxplots) <- str_replace_all(gdsc_g2p_drugboxplots, c("_gdsc_drugboxplots.png" = "", "./plots_18Q3/manuscript/gdsc_g2p_drugboxplots/" = ""))
bsselect(gdsc_g2p_drugboxplots, type = "img", live_search = TRUE, show_tick = TRUE, height = 100)
```


## G2P level A gene-drug associations only

### CCLE

```{r fig.height=4, fig.width=12}
ccle_g2p_order <- as.character(ccle_signif_g2p$Drug_Gene)
ccle_data_g2p$Drug_Gene_Ordered <- factor(ccle_data_g2p$Drug_Gene, levels = ccle_g2p_order, labels = sapply(strsplit(x = ccle_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
ccle_signif_g2p$Drug_Gene_Ordered <- factor(ccle_signif_g2p$Drug_Gene, levels = ccle_g2p_order, labels = sapply(strsplit(x = ccle_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
ccle_data_g2p_color <- as.character(ccle_data_g2p$Color_Nonsilent)
names(ccle_data_g2p_color) <- ccle_data_g2p$Mutation_Status_Nonsilent

ccle_label_text <- data.frame(p.signif.adj = ccle_signif_g2p$p.signif.adj, p.signif = ccle_signif_g2p$p.signif, Drug_Gene_Ordered = ccle_signif_g2p$Drug_Gene_Ordered)

ccle_data_g2p_plot <- ggplot(data = ccle_data_g2p, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
  facet_wrap(~ Drug_Gene_Ordered, drop = FALSE, nrow = 1) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85), outlier.shape = 3, outlier.size = 0.5) +
  scale_fill_manual(values = ccle_data_g2p_color) +
  guides(color = FALSE) +
  geom_text(data = ccle_label_text, mapping = aes(x = 1.5, y = 7, label = p.signif), nudge_y = 0.1) +
  theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "CCLE", fill = "Mutation Status", y = "AUC", x = "Mutation Status")
ccle_data_g2p_plot

# ggsave("./plots_18Q3/manuscript/ccle_data_g2p_plot.png", ccle_data_g2p_plot, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

### CTRP

```{r fig.height=7, fig.width=12}
ctrp_g2p_order <- as.character(ctrp_signif_g2p$Drug_Gene)
ctrp_data_g2p$Drug_Gene_Ordered <- factor(ctrp_data_g2p$Drug_Gene, levels = ctrp_g2p_order, labels = sapply(strsplit(x = ctrp_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
ctrp_signif_g2p$Drug_Gene_Ordered <- factor(ctrp_signif_g2p$Drug_Gene, levels = ctrp_g2p_order, labels = sapply(strsplit(x = ctrp_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
ctrp_data_g2p_color <- as.character(ctrp_data_g2p$Color_Nonsilent)
names(ctrp_data_g2p_color) <- ctrp_data_g2p$Mutation_Status_Nonsilent

ctrp_label_text <- data.frame(p.signif.adj = ctrp_signif_g2p$p.signif.adj, p.signif = ctrp_signif_g2p$p.signif, Drug_Gene_Ordered = ctrp_signif_g2p$Drug_Gene_Ordered)

ctrp_data_g2p_plot <- ggplot(data = ctrp_data_g2p, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
  facet_wrap(~ Drug_Gene_Ordered, drop = FALSE, nrow = 3) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85), outlier.shape = 3, outlier.size = 0.5) +
  scale_fill_manual(values = ctrp_data_g2p_color) +
  guides(color = FALSE) +
  geom_text(data = ctrp_label_text, mapping = aes(x = 1.5, y = 22.5, label = p.signif), nudge_y = 0.1) +
  theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "CTRP", fill = "Mutation Status", y = "AUC", x = "Mutation Status")
ctrp_data_g2p_plot

# ggsave("./plots_18Q3/manuscript/ctrp_data_g2p_plot.png", ctrp_data_g2p_plot, device = "png", dpi = 450, width = 12, height = 7, units = "in")
```

### GDSC

```{r fig.height=7, fig.width=12}
gdsc_g2p_order <- as.character(gdsc_signif_g2p$Drug_Gene)
gdsc_data_g2p$Drug_Gene_Ordered <- factor(gdsc_data_g2p$Drug_Gene, levels = gdsc_g2p_order, labels = sapply(strsplit(x = gdsc_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
gdsc_signif_g2p$Drug_Gene_Ordered <- factor(gdsc_signif_g2p$Drug_Gene, levels = gdsc_g2p_order, labels = sapply(strsplit(x = gdsc_g2p_order, split = "_"), function(x) paste0(x[1], "\n", x[2])))
gdsc_data_g2p_color <- as.character(gdsc_data_g2p$Color_Nonsilent)
names(gdsc_data_g2p_color) <- gdsc_data_g2p$Mutation_Status_Nonsilent

gdsc_label_text <- data.frame(p.signif.adj = gdsc_signif_g2p$p.signif.adj, p.signif = gdsc_signif_g2p$p.signif, Drug_Gene_Ordered = gdsc_signif_g2p$Drug_Gene_Ordered)

gdsc_data_g2p_plot <- ggplot(data = gdsc_data_g2p, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
  facet_wrap(~ Drug_Gene_Ordered, drop = FALSE, nrow = 3) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85), outlier.shape = 3, outlier.size = 0.5) +
  scale_fill_manual(values = gdsc_data_g2p_color) +
  guides(color = FALSE) +
  geom_text(data = gdsc_label_text, mapping = aes(x = 1.5, y = 1, label = p.signif), nudge_y = 0.1) +
  theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "GDSC", fill = "Mutation Status", y = "AUC", x = "Mutation Status")
gdsc_data_g2p_plot

# ggsave("./plots_18Q3/manuscript/gdsc_data_g2p_plot.png", gdsc_data_g2p_plot, device = "png", dpi = 450, width = 12, height = 7, units = "in")
```

# CRISPR: Point mutation mapping

--------------------------------------------------------------------------------

Match specific point mutations.
```{r fig.height=5, fig.width=20, eval=FALSE}
g2p_indications <- filter(read.delim("./data_munging/data_mutation_associations_appended.csv", sep = "\t", header = TRUE), Evidence.Level == "A")
maf_g2p_indications <- filter(maf_raw, Genome_Change %in% g2p_indications$MutationName)
crispr_g2p_indications <- merge(crispr_data_ptmuts, maf_g2p_indications, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
crispr_g2p_indications$Mutation_Status_Nonsilent <- ifelse(is.na(crispr_g2p_indications$Mutation_Status_Nonsilent), "Other", crispr_g2p_indications$Mutation_Status_Nonsilent)
dup_g2p_indications <- filter(crispr_g2p_indications[, c("Hugo_Symbol", "CCLE_Name", "Broad_ID")] %>% group_by(Hugo_Symbol, CCLE_Name, Broad_ID) %>% tally(), n > 1)
crispr_g2p_indications <- merge(crispr_g2p_indications, dup_g2p_indications, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"), all.x = TRUE)
crispr_g2p_indications$Genome_Change <- factor(crispr_g2p_indications$Genome_Change, levels = unique(crispr_g2p_indications$Genome_Change))
crispr_g2p_indications_nonNA <- filter(crispr_g2p_indications, is.na(n))

crispr_g2p_indications_plot <- ggplot(crispr_g2p_indications, aes(x = Hugo_Symbol, y = Score)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(alpha = 0.5, color = "lightgray", outlier.shape = NA) +
  geom_jitter(width = 0.3, mapping = aes(color = Mutation_Status_Nonsilent), size = 1) +
  theme_light() +
  theme(legend.title = element_blank()) +
  labs(x = "Gene", y = "CERES Score")
crispr_g2p_indications_plot

crispr_g2p_indications_muts_plot <- ggplot(crispr_g2p_indications, aes(x = Protein_Change, y = Score)) +
  facet_grid(~ Hugo_Symbol, scales = "free_x", space = "free") +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(alpha = 0.5, color = "lightgray", outlier.shape = NA) +
  geom_jitter(width = 0.1, mapping = aes(color = Mutation_Status_Nonsilent), size = 1) +
  theme_light() +
  theme(legend.title = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Gene", y = "CERES Score")
crispr_g2p_indications_muts_plot

# ggsave("./plots_18Q3/manuscript/crispr_g2p_indications_plot_test.png", crispr_g2p_indications_plot, width = 12, height = 5, units = "in")
# ggsave("./plots_18Q3/manuscript/crispr_g2p_indications_muts_plot_test.png", crispr_g2p_indications_muts_plot, width = 20, height = 5, units = "in")
```



# References

--------------------------------------------------------------------------------

Barretina, J., Caponigro, G., Stransky, N., Venkatesan, K., Margolin, A. A., Kim, S.,  Garraway, L. A. (2012). The Cancer Cell Line Encyclopedia enables predictive modelling of anticancer drug sensitivity. Nature, 483(7391), 603607. https://doi.org/10.1038/nature11003

Broad Institute Cancer Dependency Map; Cancer Data Science (2018): Cancer Dependency Map, CRISPR Avana dataset 18Q3 (Avana_public_18Q3). figshare. Fileset. doi:10.6084/m9.figshare.6931364.v1

Consortium, T. C. C. L. E., & Consortium, T. G. of D. S. in C. (2015). Pharmacogenomic agreement between two cancer cell line data sets. Nature, 528(7580), 8487. https://doi.org/10.1038/nature15736

Data Science, Cancer (2018): DEMETER2 data. figshare. Fileset. doi:10.6084/m9.figshare.6025238.v2

Doench, J. G., Fusi, N., Sullender, M., Hegde, M., Vaimberg, E. W., Donovan, K. F.,  Root, D. E. (2016). Optimized sgRNA design to maximize activity and minimize off-target effects of CRISPR-Cas9. Nature Biotechnology, 34(2), 184191. https://doi.org/10.1038/nbt.3437

Meyers, R. M., Bryan, J. G., McFarland, J. M., Weir, B. A., Sizemore, A. E., Xu, H.,  Tsherniak, A. (2017). Computational correction of copy-number effect improves specificity of CRISPR-Cas9 essentiality screens in cancer cells. Nature Genetics, 49(12), 17791784. https://doi.org/10.1038/ng.3984

McFarland, J. M., Ho, Z. V., Kugener, G., Dempster, J. M., Montgomery, P. G., Bryan, J. G.,  Tsherniak, A. (2018). Improved estimation of cancer dependencies from large-scale RNAi screens using model-based normalization and data integration. https://doi.org/10.1101/305656


# Session information

--------------------------------------------------------------------------------

```{r}
print(sessionInfo())
```

